Worflow

### 2017-06-12 R/workflow.Rmd
biocparallel for mac and linux

```{r options, echo=FALSE, results="hide",message=FALSE, error=FALSE, include=FALSE, autodep=TRUE}
knitr::opts_chunk$set(fig.align="center", cache=TRUE, error=FALSE, message=FALSE, warning=TRUE)
#bioconductor:
library(scRNAseq)
library(scone)
library(clusterExperiment)
#github
library(zinbwave)
library(slingshot)
#CRAN
library(ggplot2)
library(gplots)
library(magrittr)
library(matrixStats)
library(Rtsne)
library(RColorBrewer)
library(digest)
library(rARPACK)
set.seed(20)
if(packageVersion("clusterExperiment")<'1.3.0.9008') stop("must have current develop version to avoid bugs")
if(packageVersion("zinbwave")<'0.99.4.1') stop("must have current develop version to avoid bugs")
```

@@

```{r}
col_batch = rep(brewer.pal(9, "Set1"), 2)
names(col_batch) = unique(batch)
table(batch)
```

@@

```{r}
col_clus <- c("transparent", brewer.pal(12, "Set3"), brewer.pal(8, "Set2"))
col_clus <- col_clus[1:length(unique(clus.labels))]
names(col_clus) <- sort(unique(clus.labels))
table(clus.labels)
```

@@

```{r}
table(data.frame(batch = as.vector(batch),
                 cluster = clus.labels))
```

@@

```{r}
se <- SummarizedExperiment(assays = list(rawCounts = core),
                           colData = metadata)
```

@@

```{r}
fn <- '../data/zinb_batch.rda'
if (runZinb & !file.exists(fn)){
  mod <- model.matrix( ~ batch)
  print(system.time(zinb <- zinbDimRed(se, K = 50, X = mod, residuals = T,
                                       normalizedValues = T)))
  save(zinb, file = fn)
}else{
  load(fn)
}
```

@@

```{r}
res <- assays(zinb)[['normalizedValues']]
res[1:3,1:3]
```

@@

```{r}
res_order <- res[, order(as.numeric(batch))]
col_order <- as.numeric(batch)[order(as.numeric(batch))]
boxplot(res_order, main='Boxplot of normalized values\ncolor=batch',
        col = col_order, staplewex = 0, outline = 0, border = col_order, xaxt = 'n')
```

@@

```{r}
pca <- prcomp(t(res))
par(mfrow = c(1,2))
plot(pca$x, col = col_batch[batch], pch = 20,
     main="PCA of normalized values\ncolor=batch")
plot(pca$x, col = col_clus[as.character(clus.labels)], pch = 20,
     main = "PCA of normalized values\ncolor=cluster")
par(mfrow = c(1,1))
```

@@

```{r}
W <- colData(se)[, grepl('^W', colnames(colData(se)))]
W <- as.matrix(W)
d <- dist(W)
fit <- cmdscale(d, eig = TRUE, k = 2)
plot(fit$points, col = col_clus[as.character(clus.labels)], main = 'MDS', pch = 20,
     xlab = 'Component 1', ylab = 'Component 2')
legend(x = 'bottomright', legend = unique(names(col_clus)), cex = .5,
       fill = unique(col_clus), title = 'Sample')
```

@@

```{r}
table(primaryClusterNamed(ceObj))
sum(primaryCluster(ceObj) == -1)
```

@@

```{r}
# sampleData <- data.frame(ours = primaryCluster(ceObj))
# plotHeatmap(assays(se)$normalizedValues,
#             main = 'Normalized values, 1000 most variable genes',
#             clusterSamplesData = ceObj@dendro_samples,
#             sampleData = as.matrix(sampleData),clusterLegend=ceObj@clusterLegend[1])
# easier fix:
origClusterColors<-bigPalette[1:nlevels(colData(ceObj)$clusterLabels)]
experimentColors<-bigPalette[1:nlevels(colData(ceObj)$Experiment)]
batchColors<-bigPalette[1:nlevels(colData(ceObj)$Batch)]
metaColors<-list("Experiment"=experimentColors,"Batch"=batchColors,"clusterLabels"=origClusterColors)

@@

```{r}
palDF <- ceObj@clusterLegend[[1]]
pal <- palDF[, 'color']
names(pal) <- palDF[, 'name']
pal["-1"] = "transparent"
plot(fit$points, col = pal[primaryClusterNamed(ceObj)],
     main = 'MDS W, color = our new clusters', pch = 20,
     xlab = 'Component1', ylab = 'Component2')
legend(x = 'bottomright', legend = names(pal), cex = .5,
       fill = pal, title = 'Sample')
```

@@

```{r}
table(data.frame(original = clus.labels, ours = primaryClusterNamed(ceObj)))
```

@@

```{r}
Kvec <- c(3, 4, 5)
```

@@

```{r slingshot_unsup}
our_cl <- primaryClusterNamed(ceObj)
cl = our_cl[our_cl != "-1"]
pal = pal[names(pal) != '-1']
```

```{r}
for (k in Kvec){
  X <- W[our_cl != "-1", 1:k]

@@

```{r}
de <- read.csv('../data/oe_markers.txt', stringsAsFactors = F, header = F)
de <- de$V1
plotHeatmap(ceObj, 
            visualizeData = assays(se)$normalizedValues[rownames(se) %in% de, ],
            clusterSamplesData = "dendrogramValue",
            whichClusters = "primary",
            main = 'Normalized values, 1000 most variable genes',
            breaks = 0.99)
```

>>>>

```{r options, echo=FALSE, results="hide",message=FALSE, error=FALSE, include=FALSE, autodep=TRUE}
knitr::opts_chunk$set(fig.align="center", cache=TRUE, error=FALSE, message=FALSE, warning=TRUE)
#bioconductor:
library(scRNAseq)
library(scone)
library(clusterExperiment)
library(BiocParallel)
#github
library(zinbwave)
library(slingshot)
#CRAN
library(doParallel)
library(ggplot2)
library(gplots)
library(magrittr)
library(matrixStats)
library(Rtsne)
library(RColorBrewer)
library(digest)
library(rARPACK)
set.seed(20)
if(packageVersion("clusterExperiment")<'1.3.0.9008') stop("must have current develop version to avoid bugs")
if(packageVersion("zinbwave")<'0.99.4.2') stop("must have current develop version to avoid bugs")
```

FP: for zinbwave, parallel computing is now handle by BiocParallel. If you have a Windows machine, please update the code to allow paralell computing.

```{r parallel}
mysystem = Sys.info()[['sysname']]
switch(mysystem,
       Windows = {print("I'm a Windows PC and you should choose the parallel computing you want.")},
       Linux = {print("I'm a penguin and I'll use package MulticoreParam for parallel computing.")},
       Darwin = {print("I'm a Mac and I'll use package doParallel for parallel computing.")})

if (mysystem == 'Darwin'){
  registerDoParallel(NCORES)
  register(DoparParam())
}else if (mysystem == 'Linux'){
  register(bpstart(MulticoreParam(workers=NCORES)))
}else{
  print('Please change this to allow parallel computing')
  register(SerialParam())
}
```

@@

```{r batch}
col_batch = rep(brewer.pal(9, "Set1"), 2)
names(col_batch) = unique(batch)
table(batch)
```

@@

```{r original}
col_clus <- c("transparent", brewer.pal(12, "Set3"), brewer.pal(8, "Set2"))
col_clus <- col_clus[1:length(unique(clus.labels))]
names(col_clus) <- sort(unique(clus.labels))
table(clus.labels)
```

@@

```{r clustbatch}
table(data.frame(batch = as.vector(batch),
                 cluster = clus.labels))
```

@@

```{r se}
se <- SummarizedExperiment(assays = list(rawCounts = core),
                           colData = metadata)
```

@@

```{r zinb}
fn <- '../data/zinb_batch.rda'
if (runZinb & !file.exists(fn)){
  print(system.time(zinb <- zinbDimRed(se, K = 50, X = '~ batch',
                                       residuals = T,
                                       normalizedValues = T)))
  save(zinb, file = fn)
}else{
  load(fn)
}
```

@@

```{r norm}
res <- assays(zinb)[['normalizedValues']]
res[1:3,1:3]
```

@@

```{r lookatnorm}
res_order <- res[, order(as.numeric(batch))]
col_order <- as.numeric(batch)[order(as.numeric(batch))]
boxplot(res_order, main='Boxplot of normalized values\ncolor=batch',
        col = col_order, staplewex = 0, outline = 0, border = col_order, xaxt = 'n')
```

@@

```{r pcanorm}
pca <- prcomp(t(res))
par(mfrow = c(1,2))
plot(pca$x, col = col_batch[batch], pch = 20,
     main="PCA of normalized values\ncolor=batch")
plot(pca$x, col = col_clus[as.character(clus.labels)], pch = 20,
     main = "PCA of normalized values\ncolor=cluster")
par(mfrow = c(1,1))
```

@@

```{r lookatdimred}
W <- colData(se)[, grepl('^W', colnames(colData(se)))]
W <- as.matrix(W)
d <- dist(W)
fit <- cmdscale(d, eig = TRUE, k = 2)
plot(fit$points, col = col_clus[as.character(clus.labels)], main = 'MDS', pch = 20,
     xlab = 'Component 1', ylab = 'Component 2')
legend(x = 'bottomright', legend = unique(names(col_clus)), cex = .5,
       fill = unique(col_clus), title = 'Sample')
```

@@

```{r tableclust}
table(primaryClusterNamed(ceObj))
sum(primaryCluster(ceObj) == -1)
```

@@

```{r heatmaps}
# sampleData <- data.frame(ours = primaryCluster(ceObj))
# plotHeatmap(assays(se)$normalizedValues,
#             main = 'Normalized values, 1000 most variable genes',
#             clusterSamplesData = ceObj@dendro_samples,
#             sampleData = as.matrix(sampleData),clusterLegend=ceObj@clusterLegend[1])
# easier fix:
origClusterColors<-bigPalette[1:nlevels(colData(ceObj)$clusterLabels)]
experimentColors<-bigPalette[1:nlevels(colData(ceObj)$Experiment)]
batchColors<-bigPalette[1:nlevels(colData(ceObj)$Batch)]
metaColors<-list("Experiment"=experimentColors,"Batch"=batchColors,"clusterLabels"=origClusterColors)

@@

```{r compare}
palDF <- ceObj@clusterLegend[[1]]
pal <- palDF[, 'color']
names(pal) <- palDF[, 'name']
pal["-1"] = "transparent"
plot(fit$points, col = pal[primaryClusterNamed(ceObj)],
     main = 'MDS W, color = our new clusters', pch = 20,
     xlab = 'Component1', ylab = 'Component2')
legend(x = 'bottomright', legend = names(pal), cex = .5,
       fill = pal, title = 'Sample')
```

@@

```{r tabagain}
table(data.frame(original = clus.labels, ours = primaryClusterNamed(ceObj)))
```

@@

```{r kvec}
Kvec <- c(3, 4, 5)
```

@@

```{r removeunassigned}
our_cl <- primaryClusterNamed(ceObj)
cl = our_cl[our_cl != "-1"]
pal = pal[names(pal) != '-1']
```

```{r slingshot_unsup}
for (k in Kvec){
  X <- W[our_cl != "-1", 1:k]

@@

```{r de}
de <- read.csv('../data/oe_markers.txt', stringsAsFactors = F, header = F)
de <- de$V1
plotHeatmap(ceObj, 
            visualizeData = assays(se)$normalizedValues[rownames(se) %in% de, ],
            clusterSamplesData = "dendrogramValue",
            whichClusters = "primary",
            main = 'Normalized values, 1000 most variable genes',
            breaks = 0.99)
```
